---
title: "Homework 4"
author: "Alison Hall"
date: "2/5/2020"
output: html_document
---

**1. Suppose x=1.1, a=2.2, and b=3.3**
```{r}
x <- 1.1
a <- 2.2
b <- 3.3 
```

```{r}
z <- (x)^((a^b))
print(z)

```
```{r}
z <- ((x^a)^b)
print(z)
```

```{r}
z <- (3*x^3)+(2*x^2)+(1)
print(z)
```

**2. Create vectors using the rep and seq functions**

*a.  create this vector* (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)
```{r}
a <- c(seq(1,8,by=1),seq(7,1,by=-1))
print(a)

```

*b. create this vector now*
(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)
```{r}
b <- rep(1:5,1:5)
print(b)
```

*c. finally this one!*
(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)
```{r}
c <- rep(5:1, 1:5)
print(c)
```

**3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions `(asin(), acos(), or atan()),` convert these numbers into polar coordinates (If you don’t know what polar coordinates are, read about them on the web or in your calculus textbook).**
*make two random numbers between 0 and 1*
```{r}
coordinates <- runif(2)
print(coordinates)

```
*make an empy vector to store the polarcoordinates:*
```{r}
polarcoordinates <- vector(mode="numeric", length=2)
```

```{r}
polarcoordinates[1]<-sqrt(coordinates[1]^2+coordinates[2]^2)
polarcoordinates[2]<-atan(coordinates[2]/coordinates[1])
print(polarcoordinates)
```

**4. Noah's ark**
*original queue: sheep, fox, owl, ant*
```{r}
original <- c("sheep","fox", "owl", "ant")
print(original)

```
*a)serpent arrives* 
```{r}
a <- c(original,"serpent")
print(a)
```
*b) sheep exits the ark*
```{r}
b <- c(a[-1])
print(b)
```
*c) donkey arrives to the front*
```{r}
c <- c("donkey",b)
print(c)

```

*d)serpent leaves*
```{r}
d <- c(c[-5])
print(d)
```

*e)owl leaves*
```{r}
e <- c(d[-3])
print(e)
```

*f)aphid cuts the ant in like*
```{r}
f <- append(e, "aphid",after=2)
print(f)

```

g) which position is aphid in, in the queue after f?
```{r}
aphidposition <- which(f=="aphid")
print(aphidposition)
```

**5) print a sequence of numbers from 1-100 not divisible by 2,3, or 7**
```{r}
fullset <- seq(1,100)
question5 <- fullset[fullset%%2 !=0 & fullset%%3 !=0 & fullset%%7 !=0]
print(question5)

```

