---
title: "Homework 7"
author: "Alison Hall"
date: "2/26/2020"
output: html_document
---


*Hypothesis*

+ Long acclimitization time to lowered salinity levels will allow urchins to right themselves more quickly after experiencing a sudden low salinity event
+ If this is the case we would expect that there would be a negative correlation after running a linear regression, meaning that as acclimation time decreases, righting time increases. 
+ Independent variable- acclimatin time (continuous, real number)
+Dependent variable- righting time (continuous real number)

*Generation of Fake Data assuming a normal distribution in conditions where my hypothesis was true*

```{r}
library(ggplot2)
Righting_Time <- rnorm(n=10,mean=5,sd=15)
noise <- rnorm(n=10,sd=100) #to add the noise into the relatioship so it isnt exactly linear. Changing the sd makes the data more noisy. n has to be the same in both to ensure that x and y are the same length
Acclimation_Time <- 5 - Righting_Time*15 + noise
plot(Righting_Time,Acclimation_Time)
```

*Make a tibble to organize these data*
```{r}
library(dplyr)
UrchinTibble <-tibble(Righting_Time,
                      Acclimation_Time)
```

*Regression for these data*
```{r}
regression <- lm(Acclimation_Time ~ Righting_Time, UrchinTibble)
print(regression)
summary(regression)
```

*Plot these data* 
```{r}
RegressionPlot <- ggplot(data=UrchinTibble, aes (x=Righting_Time,y=Acclimation_Time))+geom_point()+stat_smooth(method = lm)
print(RegressionPlot)
```

*Adjusting the sample size*

+ I adjusted the sample size between 10 and 100 and still observe a negative linear relationship with Pvalues below 0.05 which is what I expect should my hypothesis be correct and suggests that I could still run my experiment even with just a few sea urchins. 

*Adjusting the standard deviation*

+ I adjusted the standard deviation in the noise that was added into the independent variable and the sd of the dependent variable. I notice when I adjusted the standard deviation of the noise the linear relationship was less tight. When the standard deviation of the noise was 200+ I began to observe an insignificant P value. 






