---
title: "Homework 5"
author: "Alison Hall"
date: "2/12/2020"
output: html_document
---

##Working with matrices, lists, and data frames

**Assign to the variable n_dims a single random integer between 3 and 10.**

```{r}
n_dims <- sample(3:10, 1)
print(n_dims)
```


**Create a vector of consecutive integers from 1 to n_dims2**
```{r}
n_dims2 <- seq(1,(n_dims^2), by=1)
print(n_dims2)
```

**Use the sample function to randomly reshuffle these values**

```{r}
vector <- sample(n_dims2, replace = TRUE)
```


**create a square matrix with these elements, and print it**
```{r}
squarematrix <- matrix(data=vector, nrow=n_dims, ncol=n_dims)

print(squarematrix)

```


**find a function in r to transpose the matrix. This makes the rows columns and the columns rows**
```{r}
transposematrix <- t(squarematrix)

print(transposematrix)

```


**calculate the sum and the mean of the elements in the first row and the last row.**
```{r}
SumFirstRow <- sum(transposematrix[1,])
print(SumFirstRow)
MeanFirstRow <- mean(transposematrix[1,])
print(MeanFirstRow)
SumLastRow <- sum(transposematrix[n_dims,])
print(SumLastRow)
MeanLastRow <- mean(transposematrix[n_dims,])
print(MeanLastRow)
```


**read about the eigen() function and use it on your matrix**
```{r}
eigen(transposematrix)

```


**look carefully at the elements of `$values` and `$vectors`. What kind of numbers are these?**

*Eigen vectors are non zero vectors that change by a scaler vector. The resulting value after that scaling takes place is called the eigen value.*

**dig in with the typeof() function to figure out their type.**
```{r}
typeof(eigen(transposematrix)$values)
typeof(eigen(transposematrix)$vectors)
```

**Create a list with the following named elements:**

* my_matrix, which is a 4 x 4 matrix filled with random uniform values

* my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.

* my_letters, which is a 26-element vector of all the lower-case letters in random order.

*components of the list*
```{r}
my_matrix <- matrix(runif(4),nrow=4, ncol = 4)
randomvaluesec <- runif(100)
my_logical <- (randomvaluesec > 0.3)
my_letters <- sample(letters)
```

*turn those into a list*
```{r}
list(my_matrix,my_logical,my_letters)
```

*create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.*
```{r}
newlist <- list(my_matrix[2,2], my_logical[2], my_letters[2])

print(newlist)
```


* use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(newlist[[1]])
typeof(newlist[[2]])
typeof(newlist[[3]])
```
*another way to see the type of elements in a list is this*
```{r}
str(newlist)
```


* combine the underlying elements from the new list into a single atomic vector with the c() function.
what is the data type of this vector?

```{r}
list <- c(typeof(newlist[[1]]), typeof(newlist[[2]]), typeof(newlist[[3]]))
print(list)
typeof(list)
```


**Create a data frame with two variables (= columns) and 26 cases (= rows)**


*call the first variable my_unis and fill it with a random uniform value from 0 to 10

*call the second variable my_letters and fill it with 26 capital letters in random order.


```{r}
my_units <- runif(26, min=0, max = 10)
my_letters <- sample(LETTERS)
my_dataframe <- data.frame(my_units,my_letters)

```
*for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA
```{r}
my_dataframe[sample(nrow(my_dataframe), 4), 1] <- NA
print(my_dataframe)

```


*for the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
print(which(is.na(my_dataframe)))
```



*for the second variable, sort it in alphabetical order
```{r}
my_dataframe[order(my_dataframe$my_letters),]
print(my_dataframe)
```


*calculate the column mean for the first variable.
```{r}
mean(my_dataframe[["my_units"]], na.rm = TRUE)
```

